---
title: "Using A Multivariate Model to Understand How Youths’ In-The-Moment Engagement Predicts Changes in Youths’ Interest"
author: "Joshua M. Rosenberg, Patrick N. Beymer, Thomas M. Houslay, and Jennifer A. Schmidt"
date: "3/27/2018"
output: html_document
---

```{r, setup-results-fix, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      collapse = TRUE,
                      fig.width = 6,
                      fig.asp = .618,
                      out.width = "80%", 
                      fig.align = "center", 
                      results = "hold", 
                      cache = TRUE) 

options(knitr.kable.NA = '')
set.seed(12345)
```

## Loading data, setting up

```{r, load-packages-load-data}
library(tidyverse)
library(brms)

d_red <- read_csv("processed-data/data-to-model.csv")
```

*Note to Tom*: Using results of `r2glmm::r2beta()` to calculate correlation between RM engagement (BLUP) and post-interestinterest; take square root of that value.

${ r }_{ BLUP-post-interest }$ = `r partial_corr`, accounting for their initial interest in STEM, their gender, and the nesting and cross-classification of youths’ responses.

### Super high-level summary of results from lme4

RM engagement (w/ BLUP) seems important; pre-interest seems important; gender female not quite important.

## 2B. Results using MCMCglmm

```{r, MCMC-model, eval = FALSE}
d_for_m2 <- filter(d_red, !is.na(gender_female) & !is.na(pre_interest)) # # if there are missing vals in the fixed predictors, MCMCglmm gives a warning

d_for_m2 <- fill(d_for_m2, pre_interest, post_interest)

# brms - not run

library(brms)
m3 <- brm(mvbind(rm_engagement, post_interest) ~ 
            gender_female + 
            pre_interest +
            (1|p|participant_ID) +
            (1|q|program_ID),
          data = d_for_m2, chains = 4, cores = 4, iter = 1000,
          control = list(adapt_delta = .999,
                         max_treedepth = 15))

saveRDS(m3, "m3.rds")
```

```{r}
m <- read_rds("m3.rds")
m
```