---
title: "MCMCglmm test"
author: "Joshua Rosenberg"
date: "3/2/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE)
```

The aim of this analysis is to carry out in a single model (using MCMCglmm) what I have been running in separate models. This document presents my attempt to run three models:

1. An example from a thread with Tom Houslay/Jarred Hadfield in response to a question to R-sig-mixed-models (https://stat.ethz.ch/pipermail/r-sig-mixed-models/2017q4/026092.html)
2. My attempt to do the analysis using two separate models using lme4/ best linear unbiased predictors (BLUPs)
3. My attempt to fit a similar model as in step 2, but using MCMCglmm

** Loading packages** 

```{r}
library(MCMCglmm)
library(lme4)
library(tidyverse)
```


```{r, setup-results-fix, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE, 
                      warning = FALSE, 
                      collapse = TRUE,
                      error = TRUE,
                      fig.width = 6,
                      fig.asp = .618,
                      out.width = "80%", 
                      fig.align = "center", 
                      results = "hold") 

options(knitr.kable.NA = '')
```

```{r, loading-packages}
library(tidyverse)
library(lme4)
library(corrr)
library(jmRtools)
library(tidyLPA)
library(kableExtra)
library(sjPlot)
library(broom)
library(broom.mixed)
library(konfound)
```

```{r, loading-data, eval = F}
esm <- read_csv("/Volumes/SCHMIDTLAB/PSE/data/STEM-IE/STEM-IE-esm.csv")
pre_survey_data_processed <- read_csv("/Volumes/SCHMIDTLAB/PSE/data/STEM-IE/STEM-IE-pre-survey.csv")
post_survey_data_partially_processed <- read_csv("/Volumes/SCHMIDTLAB/PSE/data/STEM-IE/STEM-IE-post-survey.csv")
video <- read_csv("/Volumes/SCHMIDTLAB/PSE/data/STEM-IE/STEM-IE-video.csv")
pqa <- read_csv("/Volumes/SCHMIDTLAB/PSE/data/STEM-IE/STEM-IE-pqa.csv")
attendance <- read_csv("/Volumes/SCHMIDTLAB/PSE/data/STEM-IE/STEM-IE-attendance.csv")
class_data <- read_csv("/Volumes/SCHMIDTLAB/PSE/data/STEM-IE/STEM-IE-class-video.csv")
demographics <- read_csv("/Volumes/SCHMIDTLAB/PSE/data/STEM-IE/STEM-IE-demographics.csv")
pm <- read_csv("/Volumes/SCHMIDTLAB/PSE/Data/STEM-IE/STEM-IE-program-match.csv")
```

```{r, loading-rdata}
# save.image("~/desktop/sandbox-01.Rdata")
load("~/desktop/sandbox-01.Rdata")
```

```{r, processing-attendance-demo-esm-data}
attendance <- rename(attendance, participant_ID = ParticipantID)
attendance <- mutate(attendance, prop_attend = DaysAttended / DaysScheduled, 
                     participant_ID = as.integer(participant_ID))
attendance <- select(attendance, participant_ID, prop_attend)

demographics <- filter(demographics, participant_ID!= 7187)
demographics <- left_join(demographics, attendance)

esm$overall_engagement <- jmRtools::composite_mean_maker(esm, hard_working, concentrating, enjoy, interest)
```

```{r, joining-to-df}
df <- left_join(esm, pre_survey_data_processed, by = "participant_ID") # df & post-survey
df <- left_join(df, video, by = c("program_ID", "response_date", "sociedad_class", "signal_number")) # df & video
df <- left_join(df, demographics, by = c("participant_ID", "program_ID")) # df and demographics
```

```{r, proc-beep-actvariables, echo = F}
df$participant_ID <- as.integer(df$participant_ID)
df$program_ID <- as.integer(df$program_ID)
df$beep_ID <- as.factor(df$beep_ID)
df$beep_ID_new <- as.factor(df$beep_ID_new)

df$youth_activity_rc <- ifelse(df$youth_activity == "Off Task", "Not Focused", df$youth_activity)

df$youth_activity_rc <- ifelse(df$youth_activity_rc == "Student Presentation" | df$youth_activity_rc == "Problem Solving", "Creating Product", df$youth_activity_rc)

df$youth_activity_rc <- ifelse(df$youth_activity_rc == "Showing Video", "Program Staff Led", df$youth_activity_rc)

df$youth_activity_rc <- as.factor(df$youth_activity_rc)

df$youth_activity_rc <- forcats::fct_relevel(df$youth_activity_rc, "Not Focused")

df$relevance <- jmRtools::composite_mean_maker(df, use_outside, future_goals, important)
```

```{r proc-demographics}
df$urm <- ifelse(df$race %in% c("White", "Asian"), 0, 1)
df$race <- as.factor(df$race)
df$race <- fct_lump(df$race, n = 2)
df$race_other <- fct_relevel(df$race, "Other")
df$gender_female <- as.factor(df$gender) # female is comparison_group
df$gender_female <- ifelse(df$gender_female == "F", 1, 
                           ifelse(df$gender_female == "M", 0, NA))
```

# josh new example

```{r}
ps <- read_csv("~/Google Drive/1_Research/dissertation/rosenberg-diss/STEM-IE-post-survey.csv")
dff <- select(df, participant_ID, program_ID, rm_engagement = overall_engagement, gender_female, urm, pre_interest = overall_pre_interest)
dff$participant_ID <- as.integer(dff$participant_ID)
ps <- select(ps, participant_ID, overall_post_interest)

d <- left_join(dff, ps, by = "participant_ID")
d$program_ID <- as.integer(d$program_ID)
d <- rename(d, post_interest = overall_post_interest)

# Preparing/structuring data 
ind <- distinct(d, participant_ID, post_interest)
rep <- select(d, participant_ID, rm_engagement)
type <- c(rep("s", nrow(ind)), rep("r", nrow(rep)))
dd <- data.frame(y = c(ind$post_interest, rep$rm_engagement),
                 type = as.factor(type),
                 individual = as.factor(c(as.character(ind$participant_ID), as.character(rep$participant_ID))))

d_red <- d %>% 
    group_by(participant_ID, program_ID) %>% 
    mutate(rownum = row_number()) %>% 
    mutate(post_interest = ifelse(rownum == 1, post_interest, NA))
```

```{r, m1}
prior1 =list(R =list(V =diag(c(1,0.0001),2,2), nu = 0.002, fix = 2),
             G =list(G1 =list(V =diag(2), nu = 2,
                              alpha.mu =rep(0,2),
                              alpha.V =diag(25^2,2,2)),
                     G2 =list(V =diag(2), nu = 2,
                              alpha.mu =rep(0,2),
                              alpha.V =diag(25^2,2,2))))

m1 <- MCMCglmm(cbind(rm_engagement, post_interest) ~ trait -1,
               random=~us(trait):participant_ID +
                   us(trait):program_ID, 
               rcov=~idh(trait):units, # this denotes response vars varying across response rows
               family = rep("gaussian",2),
               data=as.data.frame(d_red),
               prior=prior1,
               verbose=FALSE)

summary(m1)

m1_cor <- m1$VCV[,"traitpost_interest:traitrm_engagement.participant_ID"]/
    (sqrt(m1$VCV[,"traitpost_interest:traitpost_interest.participant_ID"])*
         sqrt(m1$VCV[,"traitrm_engagement:traitrm_engagement.participant_ID"]))

plot(m1_cor)
HPDinterval(m.test4_cor)
posterior.mode(m.test4_cor)
```

```{r, m2}
prior2 =list(R =list(V =diag(c(1,0.0001),2,2), nu = 0.002, fix = 2),
             G =list(G1 =list(V =diag(2), nu = 2,
                              alpha.mu =rep(0,2),
                              alpha.V =diag(25^2,2,2)),
                     G2 =list(V =diag(2), nu = 2,
                              alpha.mu =rep(0,2),
                              alpha.V =diag(25^2,2,2))))

m2 <- MCMCglmm(cbind(rm_engagement, post_interest) ~ trait - 1 + gender_female + urm + ,
               random=~us(trait):participant_ID +
                   us(trait):program_ID, 
               rcov=~idh(trait):units, # this denotes response vars varying across response rows
               family = rep("gaussian",2),
               data=as.data.frame(d_red),
               prior=prior1,
               verbose=FALSE)

summary(m1)

m1_cor <- m1$VCV[,"traitpost_interest:traitrm_engagement.participant_ID"]/
    (sqrt(m1$VCV[,"traitpost_interest:traitpost_interest.participant_ID"])*
         sqrt(m1$VCV[,"traitrm_engagement:traitrm_engagement.participant_ID"]))

plot(m1_cor)
HPDinterval(m.test4_cor)
posterior.mode(m.test4_cor)
```